---
title: "Netlyse: A Netflix analysis with R and Quarto"
author: "Faranak Rahimi & Nils Rechberger"
date: "January 2026"
toc: true
format: pdf
---

## Abstract

## Introduction

## Research Questions

::: {.callout}
How do production characteristics such as runtime, cast size, and content type influence the distribution of IMDb scores and ratings for Netflix titles?
:::

## Data Overview

### Data Sample

#### Dataset "Netflix Movies and TV Shows"

```{r}
#| echo: false

library(tidyverse)
```

```{r}
#| echo: false

netflix_data <- readr::read_csv(
    file = "data/netflix_titles.csv",
    quote = "\""
) %>% dplyr::select(
    c("title", "country", "duration")
)

netflix_data

```

#### Dataset "Netflix IMDB Scores"

```{r}
#| echo: false

imdb_data <- readr::read_csv(
    file = "data/Netflix TV Shows and Movies.csv",
    quote = "\""
) %>% dplyr::select(
    c("title", "type", "imdb_score")
)

imdb_data
```

#### Join data

```{r}
source("scripts/join_data.R") # Self build utility script

joined_data <- readr::read_csv(
    file = "data/joined_data.csv",
    quote = "\""
) %>% dplyr::select(
    c("title", "country", "duration", "type.y", "imdb_score")
)

joined_data
```

## Exploratory Data Analysis

## Statistical Analysis & Methodology

## Time Series Forecasting

## Discussion

## Conclusion

## References

## Appendix