---
title: "Netlyse: A Netflix analysis with R and Quarto"
author: "Faranak Rahimi & Nils Rechberger"
date: "January 2026"
toc: true
format: pdf
---

## Abstract

## Introduction


## Research Questions

::: {.callout}
How has Netflix’s content catalogue evolved over time in terms of content type and country of production?
:::

To address this main research question, the following sub-questions are investigated:

1. How has the number of Netflix titles changed over time?
2. How has the balance between movies and TV shows evolved on Netflix?
3. Which countries contribute the most content to Netflix?
4. How does the distribution of content types differ across producing countries?


## Data Overview

### Data Sample

#### Dataset "Netflix Movies and TV Shows"

```{r}
#| echo: false

library(tidyverse)
```

```{r}
#| echo: false

netflix_data <- readr::read_csv(
    file = "data/netflix_titles.csv",
    quote = "\""
) %>% dplyr::select(
    c("title", "country", "duration")
)

netflix_data

```

#### Dataset "Netflix IMDB Scores"

```{r}
#| echo: false

imdb_data <- readr::read_csv(
    file = "data/Netflix TV Shows and Movies.csv",
    quote = "\""
) %>% dplyr::select(
    c("title", "type", "imdb_score")
)

imdb_data
```

#### Join data

```{r}
source("scripts/join_data.R") # Self build utility script

joined_data <- readr::read_csv(
    file = "data/joined_data.csv",
    quote = "\""
) %>% dplyr::select(
    c("title", "country", "duration", "type.y", "imdb_score")
)

joined_data
```

## Exploratory Data Analysis

## Statistical Analysis & Methodology

## Time Series Forecasting

## Discussion

## Conclusion

## References

## Appendix