---
title: "Netlyse: A Netflix analysis with R and Quarto"
author: "Faranak Rahimi & Nils Rechberger"
date: "January 2026"
toc: true
format: pdf
---

## Abstract

## Introduction

### A brief History of Netflix

Netflix is an subscription video on-demand streaming service based in the USA. The service primarily distributes original and acquired films and television shows from various genres, and it is available internationally in multiple languages^[https://en.wikipedia.org/wiki/Netflix].

In the early 2000s, Netflix was launched as a mail-based rental business. With this new business model, Neftlix directly compete with Blockbuster, which was one of the biggest player in this time. Not restricted to the size of a retail store, Netflix could offer hundreds of thousands of DVDs to customers in a long tail business model.

But with the fast tecnical transformation in the later 2000s, Netflix needed to refocus there businnes model. By the mid-2000s data speeds and bandwidth costs improved sufficiently to allow customers to download movies from the internet instead to order them by mail.

### An Analytical View

Netflix had long closely analyzed its customers' preferences. Watch Now gave the company real-time data on their behavior, such as scenes that customers replayed or skipped, or when they stopped watching a show^[https://www.nplusonemag.com/issue-49/essays/casual-viewing/].

## Research Questions

::: {.callout}
How has Netflix’s content catalogue evolved over time in terms of content type and country of production?
:::

To address this main research question, the following sub-questions are investigated:

1. How has the number of Netflix titles changed over time?
2. How has the balance between movies and TV shows evolved on Netflix?
3. Which countries contribute the most content to Netflix?
4. How does the distribution of content types differ across producing countries?


## Data Overview

### Data Sample

#### Dataset "Netflix Movies and TV Shows"

```{r}
#| echo: false

library(tidyverse)
```

```{r}
#| echo: false

netflix_data <- readr::read_csv(
    file = "data/netflix_titles.csv",
    quote = "\""
) %>% dplyr::select(
    c("title", "country", "duration")
)

netflix_data

```

#### Dataset "Netflix IMDB Scores"

```{r}
#| echo: false

imdb_data <- readr::read_csv(
    file = "data/Netflix TV Shows and Movies.csv",
    quote = "\""
) %>% dplyr::select(
    c("title", "type", "imdb_score")
)

imdb_data
```

#### Join data

```{r}
source("scripts/join_data.R") # Self build utility script

joined_data <- readr::read_csv(
    file = "data/joined_data.csv",
    quote = "\""
) %>% dplyr::select(
    c("title", "country", "duration", "type.y", "imdb_score")
)

joined_data
```

## Exploratory Data Analysis

## Statistical Analysis & Methodology

## Time Series Forecasting

## Discussion

## Conclusion

## References

## Appendix